\bibitem{card1983}
Card, S. K., Moran, T. P., \& Newell, A. (1983). \textit{The Psychology of Human-Computer Interaction}. Lawrence Erlbaum Associates.

\bibitem{margono1987}
Margono, S. \& Shneiderman, B. (1987). A Study of File Manipulation by Novices Using Commands vs. Direct Manipulation. \textit{Proceedings of the 26th Annual Technical Symposium of the Washington D.C. Chapter of the ACM}. ACM Press / National Bureau of Standards. PDF available online.

\bibitem{hirsh1997}
Hirsh, H. \& Davison, B. D. (1997). An Adaptive UNIX Command-Line Assistant. \textit{Proceedings of the 1st International Conference on Autonomous Agents (Agents '97)}. ACM. PDF available at ACM Digital Library.

\bibitem{gharehyazie2016}
Gharehyazie, M., Zhou, B., \& Neamtiu, I. (2016). Expertise and Behavior of Unix Command Line Users: an Exploratory Study. \textit{Proceedings of the Ninth International Conference on Advances in Computer-Human Interactions (ACHI)}, 315-322.

\bibitem{wilson2016}
Wilson, G. (2016). Software Carpentry: Lessons Learned. \textit{F1000Research}, 3:62. DOI: 10.12688/f1000research.3-62.v2.

\bibitem{lin2018}
Lin, Z., Raychev, V., \& Vechev, M. (2018). NL2Bash: A Dataset and Benchmark for Natural Language to Bash Command Translation. \textit{Proceedings of the 11th International Conference on Language Resources and Evaluation (LREC)}.

\bibitem{singh2020}
Singh, T. D., Khilji, F. U. R., Abdullah, Divyansha, Singh, V., Apoorva, Thokchom, S., \& Bandyopadhyay, S. (2020). Seq2Seq and Joint Learning Based UNIX Command Line Prediction System. \textit{arXiv preprint arXiv:2006.11558}. PDF available on arXiv.

\bibitem{haro2022}
Haro, Y., Song, H., Dong, L., Huang, S., Chi, Z., Wang, W., Ma, S., \& Wei, F. (2022). Language Models are General-Purpose Interfaces. \textit{arXiv preprint arXiv:2206.06336}. PDF available on arXiv.

\bibitem{cli2023}
CLI Command Generation Using Generative AI. (2023). Technical Report, SSGMCE Shegaon. PDF available online.

\bibitem{spinellis2023}
Spinellis, D. (2023). Commands as AI Conversations. \textit{arXiv preprint arXiv:2309.06551}. PDF available on arXiv.
