I would like to express my sincere gratitude to all those who contributed to the completion of this thesis.

First and foremost, I extend my deepest appreciation to my supervisor, [Supervisor Name], for their invaluable guidance, continuous support, and expert feedback throughout this research project. Their insights and encouragement were instrumental in shaping this work.

I am grateful to the members of my examination committee for their time, constructive feedback, and thoughtful evaluation of this research.

Special thanks go to all the participants who volunteered their time for the user testing sessions. Their willingness to engage with the experimental setup and provide honest feedback was essential for the empirical validation of this research.

I would like to acknowledge the open-source community and the developers of the various AI-CLI tools that served as inspiration and reference for this work. The collaborative nature of software development continues to drive innovation in this field.

My appreciation extends to the faculty and staff of the Department of Informatics at Ionian University for providing an excellent academic environment and access to the resources necessary for this research.

Finally, I am deeply grateful to my family and friends for their unwavering support, patience, and understanding throughout the duration of this study. Their encouragement kept me motivated during the challenging phases of this work.

This thesis would not have been possible without the collective contributions of all these individuals and institutions.

\bigskip \\
Corfu, \today \\
Thanassis Gliatis

